%===================================== CHAP 1 =================================

\chapter{Introduction}
Creating efficient structures for computers to interpret information opens up many new areas for information retrieval. One method is structuring the information as a graph with vertices connected by edges (G(v, e)). In such a structure each vertex (v) can hold a small piece of information, and the edge (e) represents a relation. When information is modeled in this way, it can be called a knowledge graph (KG).\\

Graph based models allows complex relations to be modeled. Finding relations in a graph is also computationally cheaper than other models. Using standardized models such as the Resource Description Framework (RDF), which is created specifically for graph based data, has made it possible to develop efficient storage and retrieval solutions for graph data. One such solution is the framework Jena \cite{jena2} for Java.\\

Using RDF projects like YAGO (Yet Another Great Ontology) \cite{yago} have created graphs containing spatial and temporal data (spatiotemporal). This makes it possible to apply reason to the dataset, and derive new information from what is already there. This is a feature that typically separates a KG from an ontology \cite{KGDefYago}. In a graph, spatiotemporal data will be described using at least 2 vertices, one for a start date, and one for a place. When modeling time intervals, multiple vertices are needed.\\

There are few applications using RDF as an information storage system. Some of the more well known are DBPedia, YAGO, and Creative Commons. These applications often consists of a large linked dataset, or in the case of Creative Commons, it is used for embedding licenses. The applications built with RDF usually don't have much utility. A common utility built on top of RDF data sets are tools used to describe social relations, though there have been developed applications such as MusicBrainz that use RDF for relations between songs, artists, albums and other tags given to entities.\\

Creating methods that allows for easy access to information in RDF data makes it possible to create applications with more utility. Keyword and text search in RDF data is still something being developed and improved. Using frameworks such as Jena, it is possible to build solutions for keyword search, and spatiotemporal search. In this thesis (paper?) methods for efficient keyword search on spatial and temporal data will be explored.\\

% Today large collections of information is available, but for this information to be useful, efficient retrieval methods is needed. More and more information is also being generated in collections with specific uses, such as a school wiki, or an article collection for a company. All this different types of information collections need methods for retrieval.\\

% Any collection of information can be called a knowledge base. Such a collection can be structured or unstructured, and can be any size. Larger collections will require a structure to be navigable, and to make it easier to search. Depending on the type, size and use of a collection, different system should be used. For a small collection not expected to grow much a set of unordered articles might be sufficient, while a collection such as Wikipedia will need indexing, linking, categories and searching for the information to be accessible.\\

% In any set of information there will usually be some relation to other information. This relation can be in form of a link to another piece of information, relationships between real entities, such as the relation between a capitol and a country. Relations can also be more abstract, such as two entities being placed in the same category. Using the relationship between pieces of information it is possible to create a directed graph. Such a graph is known as a knowledge graph.\\

% Using the graph structure of knowledge graphs it is possible to use traditional traversal algorithms to find sets of information. This makes knowledge graphs a versatile 

% - Information have relationships
% - A user might need parts of different sources
% - Finding specific information based on relationship, using part of different sources/ sets (?)
% - often duplicate in articles 

%Traditional retrieval consists of large databases with indexes and web crawlers gathering information.
% - Why this is important\\


% - What is the contribution of this paper\\
% This paper will propose methods of integrating time and date facts in existing search methods.\\

Research questions?\\
RQ1: How can spatio-temporal data be integrated into exiting keyword query methods for RDF data.\\
RQ2: What methods can be used to create more effective queries on RDF data.\\
RQ3: How do spatial and temporal RDF query methods differ from from other query methods.\\
RQ4: What methods can be used to query RDF data based on a users keyword search.\\

\section{Previous work}
Keyword search on RDF graphs often follows a set of common strategies. One method is to find nodes containing one or more keyword, then following the edges from the nodes to explore the graph, and find subgraphs where the combined nodes contain as many keywords as possible while also spreading as little as possible. BLINKS \cite{blinks} propose such a method, in combination with indexing and cost balancing for expanding clusters of accessed nodes. A similar approach is used by the authors in \cite{Elbassuoni:2011:KSO:2063576.2063615} where each node has an associated document containing terms from the triple. When querying the keywords are matched with these documents, creating lists based on the matching keywords, and a subgraph is constructed by joining matches from different lists.\\

Another strategy for keyword search in RDF graphs is to infer triples from the query. One such query system is used in AquaLog \cite{aqualog}. This method processes the input into a triple based representation, based on a linguistic model, and then further processes the triplets into what they call ``query triples.''

% more processing power have made it possible for exploration based aproaches to be more usefull than query based....


There has been done some work keyword search on RDF graphs. This includes \cite{4812421} and \cite{Elbassuoni:2011:KSO:2063576.2063615}. None of these takes spatial or temporal data into consideration. The methods for keyword queries outlined in the papers are based on a combination of indexing the graph and traversing subgraphs. The indexing is similar in both papers, but retrieval, scoring and traversal methods varies. One method of retrieval and scoring that is used as a baseline is a breadth first search for traversal and a minimal subgraph for scoring.\\

In paper \cite{Shi:2016:TRS:2882903.2882941} some methods for indexing, searching and ranking keyword searches on spatial RDF graphs is proposed. These methods builds upon the methods outlined in \cite{4812421, Elbassuoni:2011:KSO:2063576.2063615}. The most substantial difference is the use of R-trees to index the spatial dimension of the graph. This is done so that a subgraph containing the keywords can be retrieved given a location. This method starts by finding the closest spatial vertex and uses that as the root for the subgraph. Subgraphs rooted in a spatial node requires different ranking, and there is proposed a set of rules in the paper.\\

R-trees are designed around the spatial dimension, but some work on modifications to include a temporal dimension has been done. Some research done on spatio-temporal tree structures include \cite{Tao:2003:TOS:1315451.1315519, r-tree-spatio-temporal}. Most of the spatio-temporal trees are however created to be able to query and index moving or frequently updated objects. To include a temporal dimension some differences in the tree structures are made. Most approaches builds on the R-tree, but modifies the tree in different aspects. For the purpose of this thesis a different tree structure is not needed.\\

% TPR-trees \cite{} uses some tech ... This makes them most suited to index moving objects, with as spatio-temporal data to indicate where an object is at a given time. The work done in \cite{} builds heavily on the R-tree, and makes it suited for ...


\clearpage